---
title: "Simulations Public Choice"
output: html_document
author: "Lea Krähenmann 19-609-221 & Sofía Carlota Velez Klumpp 19-617-372"
hand-in date: "28.05.2021"
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gtools)
library(tidyverse)
library(vote)
library(ddpcr)
library(STV)
```

## Creating a voting simulation 
For the simulation to work out successfully, we require two things: 1) randomly generated voter preferences and 2) a function which yields the winner given those voting preferences. 
Firstly, we consider the voter preferences. 

## Preference generator 
```{r}
rand_vect <- function(N, M, sd = 1, pos.only = TRUE) {
  rm(.Random.seed, envir=globalenv())
  vec <- rnorm(N, M/N, sd)
  if (abs(sum(vec)) < 0.01) vec <- vec + 1
  vec <- round(vec / sum(vec) * M)
  deviation <- M - sum(vec)
  for (. in seq_len(abs(deviation))) {
    vec[i] <- vec[i <- sample(N, 1)] + sign(deviation)
  }
  if (pos.only) while (any(vec < 0)) {
    negs <- vec < 0
    pos  <- vec > 0
    vec[negs][i] <- vec[negs][i <- sample(sum(negs), 1)] + 1
    vec[pos][i]  <- vec[pos ][i <- sample(sum(pos ), 1)] - 1
  }
  vec
}
```
## Simulation with average abortion budget 96 according to voters preferences, using budget only on one alternative
```{r}
random_population_1 <- function(voters = 10000){
   voting_profiles <- 
  rbind(A = c(96, 0, 0),
        B = c(0, 96, 0),
        C = c(0, 0, 96),
        D = c(0, 0, 96)) %>% 
  as.data.frame() %>% tibble()
   
choices <- rand_vect(2, voters, sd = 0.125 * voters)

pro_choice <- max(choices)

rest <- rand_vect(2, min(choices), sd = 0.05 * voters)

rest2 <- rand_vect(2, rest[1], sd = 0.03 * voters)

voters_dist <- c(pro_choice, rest[2], rest2)

pop <- 
  voting_profiles %>% 
  rename("Pro-Choice" = 1, "Pro-Life" = 2, "Status Quo" = 3) %>% 
  slice(rep(1:n(), times = voters_dist)) %>% 
  as.data.frame

return(pop)
}

election <- function(n_voters = 10000){
  
pop <- random_population_1(voters = n_voters)

winner <- 
  pop %>% 
  tibble %>% 
  summarise_all(~sum(.x)) %>% 
  pivot_longer(everything()) %>% 
  filter(value == max(value)) %>% 
  pull(name)

return(c("Winner" = winner))
}
```

```{r}
df1 = data.frame("Winner" = "")

for(i in 1:10000){
  df1 = rbind(df1, election(10000))
  print(str_c(i, "/10000"))
}

df1 = df1[-1,] %>% tibble() %>% rename(Winner = 1)

df1

df1 %>% 
  count(Winner)
```

## Simulation with average abortion budget according to preferences, also voting for second choice alternatives
```{r}
random_population_2 <- function(voters = 10000){
   voting_profiles <- 
  rbind(A = c(80, 0, 16),
        B = c(0, 80, 16),
        C = c(0, 16, 80),
        D = c(16, 0, 80)) %>% 
  as.data.frame() %>% tibble()
   
choices <- rand_vect(2, voters, sd = 0.125 * voters)

pro_choice <- max(choices)

rest <- rand_vect(2, min(choices), sd = 0.05 * voters)

rest2 <- rand_vect(2, rest[1], sd = 0.03 * voters)

voters_dist <- c(pro_choice, rest[2], rest2)

pop <- 
  voting_profiles %>% 
  rename("Pro-Choice" = 1, "Pro-Life" = 2, "Status Quo" = 3) %>% 
  slice(rep(1:n(), times = voters_dist)) %>% 
  as.data.frame

return(pop)
}

election <- function(n_voters = 10000){
  
pop <- random_population_2(voters = n_voters)

winner <- 
  pop %>% 
  tibble %>% 
  summarise_all(~sum(.x)) %>% 
  pivot_longer(everything()) %>% 
  filter(value == max(value)) %>% 
  pull(name)

return(c("Winner" = winner))
}
```

```{r}
df2 = data.frame("Winner" = "")

for(i in 1:10000){
  df2 = rbind(df2, election(10000))
  print(str_c(i, "/10000"))
}

df2 = df2[-1,] %>% tibble() %>% rename(Winner = 1)

df2

df2 %>% 
  count(Winner)
```

## Simulation with average abortion budget, and more budget for pro-life due to affectedness, voting for one alternative 
```{r}
random_population_3 <- function(voters = 10000){
   voting_profiles <- 
  rbind(A = c(96, 0, 0),
        B = c(0, 144, 0),
        C = c(0, 0, 96),
        D = c(0, 0,96)) %>% 
  as.data.frame() %>% tibble()
   
choices <- rand_vect(2, voters, sd = 0.125 * voters)

pro_choice <- max(choices)

rest <- rand_vect(2, min(choices), sd = 0.05 * voters)

rest2 <- rand_vect(2, rest[1], sd = 0.03 * voters)

voters_dist <- c(pro_choice, rest[2], rest2)
pop <- 
  voting_profiles %>% 
  rename("Pro-Choice" = 1, "Pro-Life" = 2, "Status Quo" = 3) %>% 
  slice(rep(1:n(), times = voters_dist)) %>% 
  as.data.frame

return(pop)
}

election <- function(n_voters = 10000){
  
pop <- random_population_3(voters = n_voters)

winner <- 
  pop %>% 
  tibble %>% 
  summarise_all(~sum(.x)) %>% 
  pivot_longer(everything()) %>% 
  filter(value == max(value)) %>% 
  pull(name)

return(c("Winner" = winner))
}
```

```{r}
df3 = data.frame("Winner" = "")

for(i in 1:10000){
  df3 = rbind(df3, election(10000))
  print(str_c(i, "/10000"))
}

df3 = df3[-1,] %>% tibble() %>% rename(Winner = 1)

df3

df3 %>% 
  count(Winner)
```

## Simulation with average abortion budget, more budget for pro-life, and voting for 2 alternatives 
```{r}
random_population_4 <- function(voters = 10000){
   voting_profiles <- 
  rbind(A = c(80, 0, 16),
        B = c(0, 120, 24),
        C = c(0, 16, 80),
        D = c(16, 0, 80)) %>% 
  as.data.frame() %>% tibble()
   
choices <- rand_vect(2, voters, sd = 0.125 * voters)

pro_choice <- max(choices)

rest <- rand_vect(2, min(choices), sd = 0.05 * voters)

rest2 <- rand_vect(2, rest[1], sd = 0.03 * voters)

voters_dist <- c(pro_choice, rest[2], rest2)
pop <- 
  voting_profiles %>% 
  rename("Pro-Choice" = 1, "Pro-Life" = 2, "Status Quo" = 3) %>% 
  slice(rep(1:n(), times = voters_dist)) %>% 
  as.data.frame

return(pop)
}

election <- function(n_voters = 10000){
  
pop <- random_population_3(voters = n_voters)

winner <- 
  pop %>% 
  tibble %>% 
  summarise_all(~sum(.x)) %>% 
  pivot_longer(everything()) %>% 
  filter(value == max(value)) %>% 
  pull(name)

return(c("Winner" = winner))
}
```

```{r}
df4 = data.frame("Winner" = "")

for(i in 1:10000){
  df4 = rbind(df4, election(10000))
  print(str_c(i, "/10000"))
}

df4 = df4[-1,] %>% tibble() %>% rename(Winner = 1)

df4

df4 %>% 
  count(Winner)
```
